---
title: 'Lab: Class One Survey Analysis by Yuxin'
output:
  html_document:
    
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(results="hide", include=TRUE,
                      echo=FALSE, warning=FALSE, message=FALSE)
```
### **INSTRUCTIONS**
####  **Complete all tasks 1-7 using reproducible coding practices (annotation of all code). Use R syntax to solve all problems (i.e. do not manually change values/variables in datasets or check them manually to make sure things are correct except as a double or triple check). Please cut and paste each problem before each R markdown chunk and annotate your code within the chunks.**

##
##

##### 1. Import class one survey data from our Github site (https://github.com/kijohnson/Advanced-Data-Analysis) (The dataset is called 'Class 1 Survey Fall 2025 2.csv'), calling the R dataframe that you create *C1survey*.
```{r}
# read the C1survey dataset from Github
pacman::p_load(dplyr, readr)
C1survey<-read_csv("https://github.com/kijohnson/Advanced-Data-Analysis/raw/refs/heads/main/Class%201%20Survey%20Fall%202025%202.csv")
```

##
##

##### 2. Determine: **a.** the number of observations (i.e. the number of people who filled out the survey) and **b.** the number of variables in the dataframe.    
```{r}
# a and b determine number of observations and variables in C1survey dataset
dim(C1survey)
```

##
##

##### 3. **a.** Rename the column variables to something shorter and that is descriptive of what the variable is about (for example *like_dogs* for the 'Do you like dogs?' question variable) and **b.** Display that they are renamed. **Hints**: You can use either the rename_with function or the names function. The list of names for time is:

```{r, echo=TRUE}
c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three", "public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city",  "highest_educ_level")
```


```{r}

names(C1survey)[1:27] <- c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three", "public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city",  "highest_educ_level")

C1survey <- C1survey %>%
  rename_with(~ c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", 
                  "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", 
                  "hobby", "program", "specialization", "stat_software", "R_exp", 
                  "coding_comfort", "coding_length", "top_three", "public_health_interest", 
                  "fav_num", "bday", "bmonth", "country", "state", "city", 
                  "highest_educ_level"),
              .cols = 1:27)
# b. display the renamed variables
colnames(C1survey)

```
##
##

##### 4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the *C1survey* dataset. **Suggestion**: Try to ask ChatGPT for a solution if you are unsure. There is a really short answer to this but we want to see what students come up with!
```{r}

table(sapply(C1survey, class)) # solution 3 is the best solution using table and sapply functions 23 character and 4 integer

```

##
##
##### 5. **a.** Using code, check the *bday* and *bmonth* variables for any unusual or missing values (**hint**: use table function). If you find any, **b.** describe what you will do with the unusual values in a comment before or immediately following the code that does it, and **c.** after you have cleaned up any unusual values, find the median bday and month. *Note: please make sure that your bday and month variables are numeric before doing statistics
```{r}

# a. look at values for bday and bmonth
table(C1survey$bday, useNA = "always")
table(C1survey$bmonth)

# b. recode bday with non numeric values

# bday recoding method 1 use case_when and create a new variable called bday2 instead of overwriting
C1survey <- C1survey %>%
  mutate(bday2 = case_when(bday == "24-Jun" ~ "24",
                           is.na(bday) ~ NA_character_,
                           bday == "15-Nov" ~ "15",
                           bday == "August 2nd 1997" ~ "2", 
                           bday == "June 2nd" ~ "2",
                           TRUE ~ bday))
# bday recoding method 2
class(C1survey$bday) # check that it is coded as character class
C1survey <- C1survey %>%
  mutate(bday = if_else(bday == "24-Jun", "24", bday)) %>%
  mutate(bday = if_else(is.na(bday), NA, bday)) %>%
  mutate(bday = if_else(bday == "15-Nov", "15", bday)) %>%
  mutate(bday = if_else(bday == "August 2nd 1997", "2", bday)) %>%
  mutate(bday = if_else(bday == "June 2nd", "2", bday))
  
C1survey$bday # check values 



# convert to numeric for statistics
C1survey$bday<-as.numeric(C1survey$bday)
table(C1survey$bday, useNA = "always") # check values 

# bmonth recoding
class(C1survey$bmonth)
C1survey <- C1survey %>%
  mutate(bmonth = if_else(bmonth %in% c("june", "June"), "6", bmonth)) %>%
  mutate(bmonth = if_else(bmonth == "May",  "5", bmonth)) %>%
  mutate(bmonth = if_else(bmonth == "August", "8", bmonth)) %>%
  mutate(bmonth = if_else(bmonth == "December",  "12", bmonth)) %>%
  mutate(bmonth = if_else(bmonth == "January", "1", bmonth)) %>%
  mutate(bmonth = if_else(bmonth == "November",  "11", bmonth)) %>%
  mutate(bmonth = if_else(bmonth == "October", "10", bmonth)) %>%
  mutate(bmonth = if_else(bmonth == "Septemeber",  "9", bmonth)) 
  
  
C1survey$bmonth # check values 

# convert to numeric for statistics
C1survey$bmonth<-as.numeric(C1survey$bmonth)
C1survey$bmonth # check values 

# c. calculate medians for bday and bmonth 
median(C1survey$bday, na.rm=TRUE) # 18 
median(C1survey$bmonth, na.rm=TRUE) # 6
```
##
##
##### 6. **a.** Create a new variable called *bseason* that gives the season according to Northern Meteorological season in which respondents were born (winter = 12/1 to 2/29, spring = 3/1 to 5/31, summer = 6/1 to 8/31, fall = 9/1 to 11/30). **b**. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct **c**. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code. 
```{r}
# a.
# using dplyr
C1survey <- C1survey %>%
  mutate(bseason=case_when(bmonth %in% c(12,1,2) ~"Winter",
                           bmonth %in% c(3,4,5) ~"Spring",
                           bmonth %in% c(6,7,8) ~"Summer",
                           bmonth %in% c(9,10,11) ~"Fall"))
table(C1survey$bmonth, C1survey$bseason) # check variable

C1survey$bseason_f<-factor(C1survey$bseason, levels=c("Winter", "Spring", "Summer", "Fall"))


# b. 
table(C1survey$bmonth, C1survey$bseason)
table(C1survey$bmonth, C1survey$bseason_f)

# c.
tab<-addmargins(table(C1survey$bmonth, C1survey$bseason, useNA = "always"), 1)
tab
tab<-addmargins(table(C1survey$bmonth, C1survey$bseason_f, useNA = "always"), 1)
tab

# Births by season: Fall = 15 Winter = 10 Spring = 16 Summer = 16
```
##
##
##### 7. Pick your favorite variable or variables to analyze, come up with a question you want to answer with that variable or variables, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #'s).





